Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-19 chi2012.tex

We hypothesise that end-users will benefit from tools that allow them to effectively browse, use and consume heterogeneous structured data from diverse sources. In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.  We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of information sources, and the processes that they relied upon to perform such tasks.  Our results suggested that people rely on multiple and diverse sources over singular, integrated ones for a number of reasons, including coverage, reliability, and alternative perspectives.

@@

Third, we developed DataPalette, an interface that enables serendipitous ``data mixing'', negating the need for people to write bespoke code to effectively combine and compare heterogenous data.  It was principally designed to support people's data integration needs for data heterogeneity, using observations made from the previous studies.  In particular, we designed a user-centric interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.  We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision-making tasks using multiple sources.

@@

A number of research PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.  For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.  Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}. Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user manually specify such a mapping translation themselves.

>>>>

We hypothesise that end-users will benefit from tools that allow them effectively browse, use and consume heterogeneous structured data from diverse sources. In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.  We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of information sources, and the processes that they relied upon to perform such tasks.  Our results suggested that people rely on multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and for alternative perspectives.

@@

Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' negating the need for people to write bespoke code to effectively combine and compare heterogenous data.  It was principally designed to support people's data integration needs for data heterogeneity, using observations made from the previous studies.  In particular, we designed a user-centric interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.  We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision making tasks using multiple sources.

@@

A number of research PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.  For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.  Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}. Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
